Waving and swaying

### 2013-12-18 Paper/grass.tex
Introduction_almost

Rich set of dynamics are exhibited by sea grasses due to their interaction with water flow which can considerably affect the hydrodynamic conditions near the 
grass. Resulting change in hydrodynamic conditions relative to unobstructed flow can influence number of environmental 
variables and processes such as transport of sediments, contaminants, dissolved oxygen;
plant growth and biomass production (known as thigmomorphogenesis) etc. 

@@

Among many dynamics shown by flexible grasses due to their interaction with water flow, their response to steady current in form of large amplitude coherent oscillations known as
mo-nami (Ackerman and Okubo 1993) is focus of this paper. Similar phenomenon of large amplitude coherent oscillation of terrestrial canopy in wind have also been observed named as Honami 
(Inoue 1993).

@@

A key feature of flow through submerged vegetation is the existence of strong shear near top of canopy due to
different amount of drag experienced by fluid in and above the canopy, existence of strong shear is manifested by
presence of an inflection point in mean velocity near canopy top.

@@

Following the work of Raupach (1996) for the case of terrestrial grass, previous studies (Ghisalberti 2002) of monami 
have been done under assumption that the a pattern of coherent eddies are excited by hydrodynamic instability of shear layer
similar to that of free shear flow. Response of sea grasses to these coherent eddies are exhibited in form of large amplitude oscillations knows as mo-nami. While shear layer
model sucessfully predict frequency of passage of corent vortices for number of experimental observations,
several aspect of existing theory remain unclear. First, the proposed hydrodynamic instability of shear layer does not take into account influence of
drag due to grasses whreas shear layer itself is produced by drag force provided by vegetation. 

@@

Second, classical free shear flow is known to be unstable above a very small Reynolds number $\geq 5 $. On the contrary experimental observations
(Grizzle,1996) shows that an event of mo-nami is not observed untill flow speed exceed a certain threshold value. Although atmoshpheric flow through terrestrial canopy seems similar to 
aquatic flow through vegetation, atmoshpheric flow is essentially unbounded, due to difference of vertical extent of flow in terrestrial and aquatic case a mechanism similar to 
terrestrial case might not be adequate for aquatic scenario.

@@

 In this paper we extend the idea of Raupach et.al via incorporating drag due to presence of vegeation. Our caluation of linear stability analysis of flow in presence of grasses 
shows that drag from the grass blades damps out the Kelvin Helmholtz instability, normally observed in a free shear layer. We also show that an event of mo-nami is caused by instablity 
of fluid flow above vegeation rather than that of shear layer near grass. We further show that inclusion of grass predicts a threshold criteria for waving characterizedd by
Reynolds number. We confirm our analyis by comparing predictin from theory with field and experimental observations.  

@@

\section{The model}
We use a mean field average model for the coupling between the flow and the canopy. Drag force exerted by grass is approximated via continous drag filed which is taken as additional body force 
in momentum equation of fluid. 
%We propose a fully coupled model in order to comprehend dynamics of canopy and fluid flow, drag force excerted by grasses is
%taken as an additional body force in momentum equation of fluid. We approximate presence of grass via continuous field whose shape at any location is approximated as averaged 
%shape of grass at that location which can be obtained by assuming grass to be in quasi-equilibrium at any moment.
% and solving force balance and momentum balance for grass  
\begin{equation}
\rho \left(\bu_{t}+\bu.\grad\bu \right) = -\grad P+\mu\grad^{2}\bu +\mathbf{f}+\rho\mathbf{g}
\end{equation}
%\begin{equation}
% \mathbf{f}=-N\mathbf{f_{d}}
%\end{equation}
where $\mathbf{f}=-N\mathbf{f_d}$ is local average drag force per unit volume; $\mathbf{f_{d}}$ is drag force per unit length of grass; N is grass number density. 
Drag force exerted by fluid on grass depends on fluid density $\rho$, flow velocity $\bu$, which we model as $\mathbf{f_{d}}=C_N \rho\bu_{N}^{2}d\hat{n}+C_{T}\rho\bu_{T}^{2}d\hat{t}$ where 
$C_{N}$ and $C_{T}$ are normal and tangential drag coefficients respectively; $\bu_{T}$, $\bu_{N}$ are velocity vector along and normal to grass.
%\begin{equation}
% \mathbf{f_{d}}=C_N \rho\bu_{N}^{2}d\hat{n}+C_{T}\rho\bu_{T}^{2}d\hat{t}
%\end{equation}\

@@

we investigate stability of base flow profile which obtained by solving steady state solution of equation(1). Considering small perturbations $u, v, p$  associated 

@@

With these scaling along with use of stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$, equations can be simplified into a single equation.

@@

where $R_{e}= \frac{\rho U_0 H}{2\mu}$ is Reynolds number and $\bar{N_g} = C_N d N_g H/2$ is non-dimensional grass number density and  $D_{drag} = R_{e}\bar{N_{g}}$ is non-dimensional drag
which is zero outside the grass. Seeking wave solution in form $\left(u,v,\psi \right)= \left(\hat u, \hat v, \hat\phi \right)e^{ikx+\sigma}$, we get generalized orr-sommerfield equation 

@@

critical Reynolds number above which we get growing solution of perturbation, as a function of grass number density for different grass height shows following characteristics 

>>>>

Sea grasses exhibit rich set of dynamics due to their interaction with the flow of water and can considerably affect surrounding hydrodynamic conditions.
%Rich set of dynamics are exhibited by sea grasses due to their interaction with water flow which can considerably affect the hydrodynamic conditions near the 
%grass.
Resulting changes in hydrodynamic conditions relative to unobstructed flow can influence number of environmental variables and processes such as 
transport of sediments, contaminants, dissolved oxygen, plant growth and biomass production etc. 

@@

The response of flexible grass canopies to steady current in form of large amplitude coherent oscillations known as mo-nami (Ackerman and Okubo, 1993) plays a central role
in the recruitment of microscopic marine organism such as blue mussel larvae. The hydrodynamic mechanism underlying monami is the focus of this paper. 
%Among many dynamics shown by flexible grasses due to their interaction with water flow, their response to steady current in form of large amplitude coherent oscillations known as
%mo-nami (Ackerman and Okubo 1993) is focus of this paper. 
Similar phenomenon of large amplitude coherent oscillation of canopy in wind known as ho-nami (Inoue 1955) have also been observed.

@@

Although atmosphere flow through terrestrial canopies seems similar to aquatic flow through vegetation, atmospheric flows are essentially unbounded vertically. Another major
difference between atmospheric flow and aquatic flow is the considerable difference of stiffness of canopies, terrestrial canopies tends to be much more rigid than marine canopies.
The previous explanation of mo-nami invokes the existence of strong shear near the top of the canopy (Ghisalberti 2002, Raupach 1996) due to
different amounts of drag experienced by fluid in and above the canopy. Existence of strong shear is manifested by
the presence of an inflection point in the mean velocity profile near the top of the canopy.

@@

%Following the work of Raupach (1996) on terrestrial canopies, 
Existing explanation of monami have been done under assumption that the a pattern of coherent eddies are excited by hydrodynamic instability of shear layer
similar to that of free shear flow known as Kelvin-Helmholtz instability. Influence of these coherent eddies over sea grasses are exhibited by their large amplitude synchronous 
oscillations.
%Sea grasses exhibit large amplitude wave like motion in response to these eddies propagating downstream.
%are exhibited in form of large amplitude oscillations knows as mo-nami.
While shear layer model successfully predict frequency of mo-nami for number of experimental observations,
several aspect of existing theory remain unexplained. First, the proposed hydrodynamic instability of shear layer produced due to drag of 
grass does not take into account influence of drag due to grasses itself. 

@@

Second, classical free shear flow is known to be unstable above a very small Reynolds number $\leq 10 $. On the contrary observed critical Reynolds number (Grizzle 1996) for monami is much
higher $\sim 1000$
%(Grizzle,1996) shows that an event of mo-nami is not observed until flow speed exceed a certain threshold value. Although atmospheric flow through terrestrial canopy seems similar to 
%aquatic flow through vegetation, atmospheric flow is essentially unbounded, due to these difference of vertical extent of flow in two cases a mechanism similar to 
%terrestrial case might not be adequate for aquatic scenario.

@@

 In this paper we extend the idea of Raupach et.al via incorporating drag due to presence of vegetation. Our calculation of linear stability analysis of flow in presence of grasses 
shows that drag from the grass blades damps out the Kelvin Helmholtz instability. We also show instead that mo-nami is caused by instability 
of fluid flow above vegetation coupled with the flow through canopy. We further show that inclusion of drag due to grass predicts a threshold condition for waving where threshold is
characterized by Reynolds number. We confirm our analysis by comparing prediction from theory with field and experimental observations.  

@@

\section{The model}
We use a mean field average model for the coupling between the flow and the canopy. Drag force exerted by grass is approximated via continuous drag field 
which is incorporated as an additional body force $\mathbf{f}=-N\mathbf{f_d}$ in fluid momentum equation. 
%We propose a fully coupled model in order to comprehend dynamics of canopy and fluid flow, drag force excerted by grasses is
%taken as an additional body force in momentum equation of fluid. We approximate presence of grass via continuous field whose shape at any location is approximated as averaged 
%shape of grass at that location which can be obtained by assuming grass to be in quasi-equilibrium at any moment.
% and solving force balance and momentum balance for grass  
\begin{equation}
\rho \left(\bu_{t}+\bu.\grad\bu \right) = -\grad P+\mu\grad^{2}\bu +\mathbf{f}+\rho\mathbf{g}
\end{equation}
%\begin{equation}
% \mathbf{f}=-N\mathbf{f_{d}}
%\end{equation}
where $\mathbf{f_{d}}$ is drag force per unit length of grass; N is grass number density. Drag force exerted by fluid on grass depends on
fluid density $\rho$, flow velocity $\bu$, which we model as $\mathbf{f_{d}}=C_N \rho\bu_{N}^{2}d\hat{n}+C_{T}\rho\bu_{T}^{2}d\hat{t}$ where 
$C_{N}$ and $C_{T}$ are normal and tangential drag coefficients respectively; $\bu_{T}$, $\bu_{N}$ are velocity vector along and normal to grass.
%\begin{equation}
% \mathbf{f_{d}}=C_N \rho\bu_{N}^{2}d\hat{n}+C_{T}\rho\bu_{T}^{2}d\hat{t}
%\end{equation}\

@@

we investigate stability of base flow profile which is obtained by solving steady state solution of equation(1). Considering small perturbations $u, v, p$  associated 

@@

With these scaling along with the use of stream function $\psi$ with $u = \psi_{y}, v= -\psi_x$, equations can be simplified into a single equation.

@@

where $R_{e}= \frac{\rho U_0 H}{2\mu}$ is Reynolds number and $\bar{N_g} = C_N d N_g H/2$ is non-dimensional grass number density and  $D_{drag} = R_{e}\bar{N_{g}}$ is non-dimensional.
Seeking wave solution in form of $\left(u,v,\psi \right)= \left(\hat u, \hat v, \hat\phi \right)e^{ikx+\sigma}$, we get generalized orr-sommerfield equation 

@@

Critical Reynolds number above which we get growing solution of perturbation, as a function of grass number density for different grass height shows following characteristics 
